\vspace{-\baselineskip} Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-2\baselineskip}

### 2017-06-12 paper/ms.tex
nothing really

The product of a rank-$R$ semiseparable matrix with general vectors and
matrices can be computed in $\mathcal{O}(N\,R)$ operations
\citep{Vandebril:2007}.
In the context of \celeriteterm\ models, na\"ive application of these methods
will result in numerical overflow and underflow issues much like we found with
the Cholesky factorization.
To avoid these errors, we derive the following numerically stable algorithm
for computing the dot product of a \celeriteterm\ covariance matrix
\begin{eqnarray}
\bvec{y} = K\,\bvec{z}
\end{eqnarray}
using the semiseparable representation for $K$ defined in \eqalt{reparam}:
\begin{eqnarray}
f^+_{n,j} &=& \phi_{n+1,j}\,\left(f^+_{n+1,j} + \tilde{U}_{n,j}\,z_{n+1}\right) \\
f^-_{n,j} &=& \phi_{n,j}\,\left(f^-_{n-1,j} + \tilde{V}_{n-1,j}\,z_{n-1}\right) \\
y_n &=& A_{n,n}\,z_n + \sum_{j=1}^{2\,J} \left[ \tilde{V}_{n,j} f^+_{n,j} +
    \tilde U_{n-1,j} f^-_{n,j}\right] \quad,
\end{eqnarray}
where $f^+_{N,j} = 0$ and $f^-_{1,j} = 0$.
This algorithm requires two sweeps~--~$n=1, \ldots, N$ to compute $f^-$ and
$n=N,\ldots,1$ to compute $f^+$~--~with a scaling of $\mathcal{O}(N\,J)$.

>>>>

The product of a rank-$R$ semiseparable matrix with general vectors and
matrices can be computed in $\mathcal{O}(N\,R)$ operations
\citep{Vandebril:2007}.
In the context of \celeriteterm\ models, na\"ive application of these methods
will result in numerical overflow and underflow issues much like we found with
the Cholesky factorization.
To avoid these errors, we derive the following numerically stable algorithm
for computing the dot product of a \celeriteterm\ covariance matrix
\begin{eqnarray}
\bvec{y} = K\,\bvec{z}
\end{eqnarray}
using the semiseparable representation for $K$ defined in \eqalt{reparam}:
\begin{eqnarray}
f^+_{n,j} &=& \phi_{n+1,j}\,\left(f^+_{n+1,j} + \tilde{U}_{n,j}\,z_{n+1}\right) \\
f^-_{n,j} &=& \phi_{n,j}\,\left(f^-_{n-1,j} + \tilde{V}_{n-1,j}\,z_{n-1}\right) \\
y_n &=& A_{n,n}\,z_n + \sum_{j=1}^{2\,J} \left[ \tilde{V}_{n,j} f^+_{n,j} +
    \tilde U_{n-1,j} f^-_{n,j}\right] \quad,
\end{eqnarray}
where $f^+_{N,j} = 0$ and $f^-_{1,j} = 0$.
This algorithm requires two sweeps~--~$n=2,\,\ldots,\,N$ to compute $f^-$ and
$n=N-1,\,\ldots,\,1$ to compute $f^+$~--~with a scaling of $\mathcal{O}(N\,J)$.
